### 2020-09-15 paper/perfectGraphs.tex
Poprawki up to old Ch 1

The notion of perfect graphs was first introduced by Berge in 1961 \cite{CB61} and it indeed captures some of the idea of graph being ''simple'' -- in all perfect graphs the coloring problem, maximum (weighted) clique problem, and maximum (weighted) independent set problem can be solved in polynomial time \cite{grotschel1993}. Other classical NP-complete problems are still NP-complete in perfect graphs e.g. Hamiltonian path \cite{Mller1996}, maximum cut problem \cite{Bodlaender1994} or dominating set problem \cite{Dewdney81}. There are many subclasses of perfect graphs, we take a look and benchmark our implementations on some of them in \Cref{sec:experiments}

The most fundamental problem -- the problem of recognizing perfect graphs -- was open since its posing in 1961 until recently. Its solution, a polynomial algorithm recognizing perfect graphs is a union of the strong perfect graph theorem (\Cref{sec:SPGT}) stating that a graph is perfect if and only if it is Berge and an algorithm for recognizing Berge graphs in polynomial time (\Cref{sec:recognizingBerge}).

@@

Perfect graphs are interesting not only because of their theoretical properties, but they are also used in other areas of study e.g. integrality of polyhedra \cite{Chvtal1975, Chudnovsky2003}, radio channel assignment problem \cite{McDiarmid99, McDiarmid2000} and appear in the theory of Shannon capacity of a graph \cite{Lovasz1979}. Also, as pointed out in \cite{alfonsinPerfect2001, Chudnovsky2003} algorithms to solve semi-definite programs grew out of the theory of perfect graphs. We will take a look at semi-definite programs and at perfect graph's relation to Shannon capacity in \Cref{sec:ShannonCapacity}.

@@

The following is based on the paper by \citeauthor{MC05} \citetitle*[]{MC05} \cite{MC05}. We will not provide full proof of its correctness, but will aim to show the intuition behind the algorithm.

@@

We will say that $a$ can be \emph{linked onto} the triangle $\{b_1, b_2, b_3\}$ \emph{via} the paths $P_1$, $P_2$, $P_3$. Let us notice, that a pyramid is uniquely determined by its paths $P_1$, $P_2$, $P_3$.

@@

To prove that existence of $\T_2$ configuration implies that the graph is not Berge, we will need the following Roussel-Rubio lemma:

@@

\begin{algorithmic}[1]
	\mProcedure{Contains-T3}{$G$}
	\mForEach{$v_1, v_2, v_5 \in V(G)$, so that $v_1v_2$ is an edge \AND
		\lsx $v_1v_5, v_2v_5$ are nonedges \label{line:t3start}}
		\ls $Y \gets$ the set of all $\{v_1, v_2, v_5\}$-complete vertices.
		\mForEach{$X$ -- an anticomponent of $Y$}
			\ls $F' \gets$ maximal connected subset containing $v_5$, such that $v_1, v_2$ \label{line:t3n4start}
			\lsx have no neighbors in $F'$ and no vertex of $F'\setminus\{v_5\}$ is $X$-complete.
			\ls $F'' \gets$ the set of all $X$-complete vertices that have a neighbor in 
			\lsx $F'$ and are nonadjacent to all of $v_1, v_2$ and $v_5$
			\ls $F \gets F' \cup F''$ \label{line:t3mid}
			\mForEach{$v_4 \in V(G) \setminus\{v_1, v_2, v_5\}$, such that $v_4$ is adjacent to $v_1$
			\lsx and not to $v_2$ and $v_5$}
				\mIf{$v_4$ has a neighbor in $F$ and a nonneibhbor in $X$ \label{line:t3v4if}}
					\ls $v_6 \gets$ a neighbor of $v_4$ in $F$
					\mForEach{$v_3 \in V(G) \setminus\{v_1, v_2, v_4, v_5, v_6\}$}
						\mIf{$v_3$ is adjacent to $v_2, v_4, v_5$ and not adjacent to $v_1$ \lsx}
							\ls $P \gets$ a path from $v_6$ to $v_5$ with interior in $F'$
							\ls \RETURN \TRUE \Comment{$v_1, \ldots v_6, P, X$ is a $\T_3$}
						\mEndIf
					\mEndFor
				\mEndIf
			\mEndFor \label{line:t3n4end}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

@@

\begin{algorithmic}[1]
	\mProcedure{List-Possible-Near-Cleaners}{$G$}
	\ls $\mathcal{X} \gets \emptyset$
	\mForEach{relevant triple $(a, b, c)$}
		\ls $A \gets$ anticomponents of $N(\{a, b\})$
		\mIf{$c$ is $N(\{a, b\})$-complete}
			\ls $r \gets 0$
		\mElse
			\ls $r \gets$ the cardinality of the largest set of $A$ that contains 
			\lsx a nonneibhbor of $c$ 
		\mEndIf
		\ls $Y \gets \emptyset$ 
		\mForEach{$A_i \in A$}
			\mIf{$|V(A_i)| > r$}
				\ls $Y \gets Y \cup A_i$
			\mEndIf
			\mIf{$c \in A_i$}
				\ls $W \gets A_i$
			\mEndIf
		\mEndFor
		\ls $Z \gets$ the set of all $Y \cup W$-complete vertices
		\ls $X(a, b, c) \gets Y \cup Z)$ \Comment{we use $X(a, b, c)$ in the proof of correctness}
		\ls $\mathcal{X} \gets \mathcal{X} \cup X(a, b, c)$
	\mEndFor
	\ls $\mathcal{N} \gets \emptyset$
	\mForEach{edge $uv \in E(G)$}
		\ls $\mathcal{N} \gets \mathcal{N} \cup N(\{u, v\})$
	\mEndFor
	\ls $\mathcal{R} \gets \emptyset$
	\mForEach{$X_i \in X$}
		\mForEach{$N_j \in N$}
			\ls $\mathcal{R} \gets \mathcal{R} \cup (X_i \cup N_j)$
		\mEndFor
	\mEndFor
	\ls \RETURN $\mathcal{R}$
	\mEndProcedure
\end{algorithmic}

@@

\begin{algorithmic}[1]
	\mProcedure{Test-Near-Cleaner}{$G, X$}
		\mForEach{pair $x, y \in V(G)$, such that $x \neq y$}
			\mIf{there is a path between $x$ and $y$ with no internal vertices in $X$ \lsx}
				\ls $R(x, y) \gets$ the shortest path $R(x, y)$ between $x, y$ with no
				\lsx internal vertex in $X$.
				\ls $r(x, y) \gets$ the length of $R(x, y)$
			\mElse
				\ls $R(x, y) \gets \textsc{undefined}$
				\ls $r(x, y) \gets \infty$
			\mEndIf
			\mForEach{$y_1 \in V(G) \setminus X$}
				\mForEach{3-vertex path $x_1-x_2-x_3$ of $G\setminus y_1$}
					\mIf{$R(x_1, y_1), R(x_2, y_2)$ both exist}
						\ls $y_2 \gets$ the neighbor of $y_1$ in $R(x_2, y_1)$
						\mIf{$r(x_2, y_1) = r(x_1, y_1) + 1 = r(x_1, y_2) =$\lsx$min(r(x_3, y_1), r(x_3, y_2))$}
						\ls \RETURN an odd hole found
						\mEndIf
					\mEndIf
				\mEndFor
			\mEndFor
		\mEndFor
		\RETURN no odd hole found
	\mEndProcedure
\end{algorithmic}

@@

		Since $r(x_1, y_1) = r(x_2, y_1) - 1$ and since $x_1$, $x_2$ are nonadjacent it follows that $x_2$ does not belong to $R(x_1, y_1)$ and $x_1$ does not belong to $R(x_2, y_1)$. Since $r(x_3, y_1), r(x_3, y_2) \geq n (= r(x_1, y_2))$ it follows that $x_3$ does not belong to $R(x_1, y_1)$ or to $R(x_2, y_1)$, and has no neighbors in $R(x_1, y_1) \cup R(x_2, y_1)$ other than $x_1$, $x_2$. Since $r(x_1, y_2) = n$ we get that $y_2$ does not belong to $R(x_1, y_1)$.\par}%

	We claim first that the insides of paths $R(x_1, y_1)$ and $R(x_2, y_1)$ have no common vertices. For suppose that there are $2 \leq i \leq n-1$ and $2 \leq j \leq n$ that $p_i = q_j$. Then the subpaths of these two paths between $p_i$, $y_1$ are both subpaths of the shortest paths, and therefore have the same length, that is $j=i+1$. So $p_1-\ldots-p_1-q_{j+1}-\ldots-q_n$ contains a path between $x_1$, $y_2$ of length $\leq n-2$, contradicting that $r(x_1, y_2) = n$. So $R(x_1, y_1)$ and $R(x_2, y_1)$ have no common vertex except $y_1$.

@@

	Now, since $x_1$, $x_2$ are nonadjacent $i \geq 2$. But also $r(x_2, y_1) \geq n$ and so $j+n-i \geq n$, that is $j \geq i$. So we get $i=j$. Let us choose $i$ minimum, then $x_3-x_1-\ldots-p_i-q_i-\ldots-x_2-x_3$ is an odd hole, which was what we wanted.
\end{proof}

>>>>

The notion of perfect graphs was first introduced by Berge in 1961 \cite{CB61} and it indeed captures some of the idea of graph being ''simple'' -- in all perfect graphs the coloring problem, maximum (weighted) clique problem, and maximum (weighted) independent set problem can be solved in polynomial time \cite{grotschel1993}. Other classical NP-complete problems remain hard in perfect graphs e.g. Hamiltonian path \cite{Mller1996}, maximum cut problem \cite{Bodlaender1994} or dominating set problem \cite{Dewdney81}. There are many well-known subclasses of perfect graphs, we take a look and benchmark our implementations on some of them in \Cref{sec:experiments}

The most fundamental problem -- the problem of recognizing perfect graphs -- was open since its posing in 1961 until recently. Its solution, a polynomial algorithm recognizing perfect graphs, is a union of the strong perfect graph theorem (\Cref{sec:SPGT}) stating that a graph is perfect if and only if it is Berge and an algorithm for recognizing Berge graphs in polynomial time (\Cref{sec:recognizingBerge}).

@@

Perfect graphs are interesting not only because of their theoretical properties, but they are also used in other areas of study e.g. integrality of polyhedra \cite{Chvtal1975, Chudnovsky2003}, radio channel assignment problem \cite{McDiarmid99, McDiarmid2000} and appear in the theory of Shannon capacity of a graph \cite{Lovasz1979}. Also, as pointed out in \cite{alfonsinPerfect2001, Chudnovsky2003}, algorithms to solve semi-definite programs grew out of the theory of perfect graphs. We will take a look at semi-definite programs and at perfect graph's relation to Shannon capacity in \Cref{sec:ShannonCapacity}.

@@

The following is based on the paper by \citeauthor{MC05} \citetitle*[]{MC05} \cite{MC05}. We will not provide full proof of its correctness, but we will aim to show the intuition behind the algorithm.

@@

We will say that $a$ is \emph{linked onto} the triangle $\{b_1, b_2, b_3\}$ \emph{via} the paths $P_1$, $P_2$, $P_3$. Let us notice, that a pyramid is uniquely determined by its paths $P_1$, $P_2$, $P_3$.

@@

To prove that the existence of $\T_2$ configuration implies that the graph is not Berge, we will need the following lemma:

@@

\begin{algorithmic}[1]
	\mProcedure{Contains-T3}{$G$}
	\mForEach{$v_1, v_2, v_5 \in V(G)$, so that $v_1v_2$ is an edge \AND
		\lsx $v_1v_5, v_2v_5$ are nonedges \label{line:t3start}}
		\ls $Y \gets$ the set of all $\{v_1, v_2, v_5\}$-complete vertices.
		\mForEach{$X$ -- an anticomponent of $Y$}
			\ls $F' \gets$ maximal connected subset containing $v_5$, such that $v_1, v_2$ \label{line:t3n4start}
			\lsx have no neighbors in $F'$ and no vertex of $F'\setminus\{v_5\}$ is $X$-complete.
			\ls $F'' \gets$ the set of all $X$-complete vertices that have a neighbor in 
			\lsx $F'$ and are nonadjacent to all of $v_1, v_2$ and $v_5$
			\ls $F \gets F' \cup F''$ \label{line:t3mid}
			\mForEach{$v_4 \in V(G) \setminus\{v_1, v_2, v_5\}$, such that $v_4$ is adjacent to $v_1$
			\lsx and not to $v_2$ and $v_5$}
				\mIf{$v_4$ has a neighbor in $F$ and a nonneighbor in $X$ \label{line:t3v4if}}
					\ls $v_6 \gets$ a neighbor of $v_4$ in $F$
					\mForEach{$v_3 \in V(G) \setminus\{v_1, v_2, v_4, v_5, v_6\}$}
						\mIf{$v_3$ is adjacent to $v_2, v_4, v_5$ and not adjacent to $v_1$ \lsx}
							\ls $P \gets$ a path from $v_6$ to $v_5$ with interior in $F'$
							\ls \RETURN \TRUE \Comment{$v_1, \ldots v_6, P, X$ is a $\T_3$}
						\mEndIf
					\mEndFor
				\mEndIf
			\mEndFor \label{line:t3n4end}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

@@

\begin{algorithmic}[1]
	\mProcedure{List-Possible-Near-Cleaners}{$G$}
	\ls $\mathcal{X} \gets \emptyset$
	\mForEach{relevant triple $(a, b, c)$}
		\ls $A \gets$ anticomponents of $N(\{a, b\})$
		\mIf{$c$ is $N(\{a, b\})$-complete}
			\ls $r \gets 0$
		\mElse
			\ls $r \gets$ the cardinality of the largest set of $A$ that contains 
			\lsx a nonneibhbor of $c$ 
		\mEndIf
		\ls $Y \gets \emptyset$ 
		\mForEach{$A_i \in A$}
			\mIf{$|V(A_i)| > r$}
				\ls $Y \gets Y \cup A_i$
			\mEndIf
			\mIf{$c \in A_i$}
				\ls $W \gets A_i$
			\mEndIf
		\mEndFor
		\ls $Z \gets$ the set of all $(Y \cup W)$-complete vertices
		\ls $X(a, b, c) \gets Y \cup Z$ \Comment{we use $X(a, b, c)$ in the proof of correctness}
		\ls $\mathcal{X} \gets \mathcal{X} \cup X(a, b, c)$
	\mEndFor
	\ls $\mathcal{N} \gets \emptyset$
	\mForEach{edge $uv \in E(G)$}
		\ls $\mathcal{N} \gets \mathcal{N} \cup N(\{u, v\})$
	\mEndFor
	\ls $\mathcal{R} \gets \emptyset$
	\mForEach{$X_i \in \mathcal{X}$}
		\mForEach{$N_j \in \mathcal{N}$}
			\ls $\mathcal{R} \gets \mathcal{R} \cup (X_i \cup N_j)$ \label{line:listNCcalcR}
		\mEndFor
	\mEndFor
	\ls \RETURN $\mathcal{R}$
	\mEndProcedure
\end{algorithmic}

@@

\begin{algorithmic}[1]
	\mProcedure{Test-Near-Cleaner}{$G, X$}
		\mForEach{pair $x, y \in V(G)$, such that $x \neq y$}
			\mIf{there is a path between $x$ and $y$ with no internal vertices in $X$ \lsx}
				\ls $R(x, y) \gets$ the shortest path $R(x, y)$ between $x, y$ with no
				\lsx internal vertex in $X$.
				\ls $r(x, y) \gets$ the length of $R(x, y)$
			\mElse
				\ls $R(x, y) \gets \textsc{undefined}$
				\ls $r(x, y) \gets \infty$
			\mEndIf
			\mForEach{$y_1 \in V(G) \setminus X$} \label{line:cudaStart}
				\mForEach{3-vertex path $x_1-x_2-x_3$ of $G\setminus y_1$} \label{line:3vertex}
					\mIf{$R(x_1, y_1), R(x_2, y_2)$ both exist} \label{line:cudaWorkStard}
						\ls $y_2 \gets$ the neighbor of $y_1$ in $R(x_2, y_1)$
						\mIf{$r(x_2, y_1) = r(x_1, y_1) + 1 = r(x_1, y_2) =$\lsx$min(r(x_3, y_1), r(x_3, y_2))$}
						\ls \RETURN an odd hole found
						\mEndIf
					\mEndIf \label{line:cudaWorkEnd}
				\mEndFor
			\mEndFor \label{line:cudaEnd}
		\mEndFor
		\ls \RETURN no odd hole found
	\mEndProcedure
\end{algorithmic}

@@

		Since $r(x_1, y_1) = r(x_2, y_1) - 1$ and since $x_1$, $x_2$ are nonadjacent it follows that $x_2$ does not belong to $R(x_1, y_1)$ and $x_1$ does not belong to $R(x_2, y_1)$. Since $r(x_3, y_1), r(x_3, y_2) \geq n = r(x_1, y_2)$ it follows that $x_3$ does not belong to $R(x_1, y_1)$ or to $R(x_2, y_1)$, and has no neighbors in $R(x_1, y_1) \cup R(x_2, y_1)$ other than $x_1$, $x_2$. Since $r(x_1, y_2) = n$ we get that $y_2$ does not belong to $R(x_1, y_1)$.\par}%

	We claim first that the insides of paths $R(x_1, y_1)$ and $R(x_2, y_1)$ have no common vertices. For suppose that there are $2 \leq i \leq n-1$ and $2 \leq j \leq n$ that $p_i = q_j$. Then the subpaths of these two paths between $p_i$, $y_1$ are both subpaths of the shortest paths, and therefore have the same length equal to $j=i+1$. So $p_1-\ldots-p_1-q_{j+1}-\ldots-q_n$ contains a path between $x_1$, $y_2$ of length $\leq n-2$, contradicting that $r(x_1, y_2) = n$. So $R(x_1, y_1)$ and $R(x_2, y_1)$ have no common vertex except $y_1$.

@@

	Now, since $x_1$, $x_2$ are nonadjacent $i \geq 2$. But also $r(x_2, y_1) \geq n$ and so $j+n-i \geq n$, which gives us $j \geq i$. So we get $i=j$. Let us choose $i$ minimum, then $x_3-x_1-\ldots-p_i-q_i-\ldots-x_2-x_3$ is an odd hole, which was what we wanted.
\end{proof}
